Foundations of ecological and evolutionary change

### 2025-08-04 foundations_of_ecoevo.Rmd
Line 165-167 address the comment about the description of Omega, which reviewer points out can more simply be described as the sum of character values

We therefore begin with the most fundamental axioms underlying the ecology and evolution of living systems [@Rice2004; @Rice2009].
In such systems, diversity is discontinuous, in the sense that living systems are composed of discrete entities including individual organisms and groups of organisms [@Dobzhansky1970].
Our framework is general enough that entities can be anything discrete, but we will focus on each entity $i$ as an individual organism.
Individuals give rise to new individuals through birth such that $\beta_{i}$ is the number of births attributable to $i$.
Individuals are removed from the population through death such that $\delta_{i}$ is an indicator variable that takes a value of 1 (death of $i$) or 0 (persistence of $i$).
All individuals are defined by some characteristic $z_{i}$, and $\Delta z_{i}$ defines any change in $z_{i}$ from one time step $t$ to the next $t + 1$.
The total number of individuals in the population is $N$.
From this foundation, we can define $\Omega$ to be a summed characteristic across $N$ entities,

@@

This foundation of eco-evolutionary change is a statistical interaction between the demographic processes of birth and death ($\beta_{i} - \delta_{i} + 1$) and some characteristic of an individual ($z_{i} + \Delta z_{i}$).
The value of $\Omega$ is a summation of individual characteristics and takes the same units as $z$.
From eqn 1, we can derive the most fundamental equations of population ecology and evolutionary biology through an appropriate interpretation of $z$.
Under more limited interpretations of $z$, we can also interpret $\Omega$ as a metric of ecosystem function.

>>>>

We therefore begin with the most fundamental axioms underlying the ecology and evolution of living systems [@Rice2004; @Rice2009].
In such systems, diversity is discontinuous, in the sense that living systems are composed of discrete entities including individual organisms and groups of organisms [@Dobzhansky1970].
Our framework is general enough that entities can be anything discrete, but we will focus on each entity $i$ as an individual organism.
Individuals give rise to new individuals through birth such that $\beta_{i}$ is the number of births attributable to $i$.
Individuals are removed from the population through death such that $\delta_{i}$ is an indicator variable that takes a value of 1 (death of $i$) or 0 (persistence of $i$).
All individuals are defined by some characteristic $z_{i}$, and $\Delta z_{i}$ defines any change in $z_{i}$ from one time step $t$ to the next $t + 1$.
The total number of individuals in the population is $N$.
From this foundation, we can define $\Omega$ to be the sum of character values in $t+1$,

@@

This foundation of eco-evolutionary change is a statistical interaction between the demographic processes of birth and death ($\beta_{i} - \delta_{i} + 1$) and some characteristic of an individual ($z_{i} + \Delta z_{i}$).
The value of $\Omega$ takes the same units as $z$.
From eqn 1, we can derive the most fundamental equations of population ecology and evolutionary biology through an appropriate interpretation of $z$.
Under more limited interpretations of $z$, we can also interpret $\Omega$ as a metric of ecosystem function.
